Titeln på avhandlingens språk

### 2023-04-01 body/part1/chapter2-relatedwork/relatedwork.tex
wip
relwork

@@

This dissertation complements existing work on benchmarking distributed system architectures.
A rich body of literature exists with examples of such approaches for Cloud computing.
For example, the \gls{YCSB}~\cite{cooper2010benchmarking} is a data-intensive benchmark for the profiling of data centers and cloud services.
Workload characterization in datacenters leads to a similar benchmark in~\cite{jia2013characterizing}.
Both these approaches focus heavily on data processing and downlink transmission in cloud computing scenarios.

In~\cite{turner2012cmart}, the authors introduce \emph{C-MART}, a benchmark for modern web applications running on the cloud.
The benchmark utilizes modern web technologies, can be deployed automatically at a large scale, and includes generators that mimic remote clients accessing the system over the internet.\@
\cite{malawski2018benchmarking,back2018using} focus instead on serverless computing, using compute-intensive workloads to benchmark major cloud function providers such as \gls{AWS}, Microsoft Azure, and \gls{GCP}.

A number of works have tackled this challenge in the context of Fog and Edge Computing as well.
\emph{EdgeBench}, introduced in~\cite{das2018edgebench}, corresponds to a benchmarking suite for the Edge.
It emulates three different workloads, speech-to-text, image recognition, and sensor network application traffic, and specifically targets serverless deployments.\@
\cite{baurle2022comb} presents \emph{ComB}, a flexible and extensible, application-oriented benchmarking suite for Edge computing.
Its stated purpose is to assist system designers and application developers to evaluate the suitability of Edge deployments.
The framework is designed to decouple orchestration and workload, allowing for easy extension or replacement of the built-in distributed video-analytics benchmark.

\emph{IoTBench}~\cite{lee2019iotbench} is a benchmarking tool for edge-deployed \gls{IOT} applications.
It includes seven representative workloads, and allows researchers to evaluate the performance of these applications on Edge infrastructure through the analysis of computational demands and resource consumption.\@
\cite{mcchesney2019defog} presents a benchmarking suite for the comparison of deployments on the Edge, Fog, or Cloud named \emph{DeFog}.
\emph{DeFog} includes six latency-critical and bandwidth-intensive, Fog- and Edge-relevant workloads, including real-time video detection and a text-to-speech application.
In~\cite{lockhart2020scission}, the authors introduce a tool, \emph{Scission}, for the benchmarking of \glspl{DNN} partitioned across the user device, edge, and cloud continuum.
The tool can also be used to automatically determine the optimal partition of a \gls{DNN}.
Finally, \emph{OpenRTiST}~\cite{george2020openrtist} is a tool for the benchmarking and performance evaluation of Edge Computing deployments using real-time processing of a live video feed.
Video is streamed from a source Android device to a compute node where neural style-transfer~\cite{gatys2016image} is applied to it.
The transformed video is then fed back to the source to be presented, resulting in a compute-intensive, bandwidth-hungry, and latency-sensitive benchmark.

@@

\gls{GENI} is a testbed deployed across more than fifty sites across the \gls{USA}.
Originally designed for ``the development, deployment, and validation of transformative, at-scale concepts in network science, services, and security''~\cite{berman2014geni}, \gls{GENI} today combines local edge compute resources with mobile network last-hop connectivity for \gls{MEC} experimentation~\cite{gosain2017geni}.
The wireless-capable sites incorporate single-hop access to compute, storage, and network resource, making \gls{GENI} a compelling Edge Computing platform.

The Chameleon testbed is a research platform originally designed for the study and evaluation of cloud computing systems~\cite{keahey2020lessons}, but which has recently been expanded to target Edge Computing research through the \acs{CHI}@Edge~\cite{chiatedge} project.
Chameleon is a collaborative project operated by a consortium of research institutions and industry partners, and is designed to provide a flexible and reconfigurable infrastructure for researchers to experiment with and test various cloud computing technologies, algorithms, and configurations.
The testbed provides a highly reconfigurable array of hardware and software resources, including servers, storage systems, and networking equipment, able to simulate various cloud and Edge Computing scenarios.

\gls{COSMOS} is a testbed deployed in New York City (State of New York, \gls{USA}), containing over \num{200} rooftop, intermediate, and mobile nodes.
The testbeds main \emph{raison d'être} is \gls{MEC} research, with a strong emphasis on wireless networks.
The system includes \glspl{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and edge compute for core network functionality and application data processing~\cite{yu2019cosmos,raychaudhuri2020challenge}.

\gls{POWDER} is described by the authors of~\cite{breen2020powder} as a ``city-scale, remotely accessible, end-to-end software defined platform supporting a broad range of wireless and mobile related research''.
It features roughly \num{15} fixed programmable radio nodes, based on off-the-shelves \glspl{SDR}, distributed across a \SI{15}{\kilo\meter\squared} area in Salt Lake City (State of Utah, \gls{USA}).
These radio nodes are combined with edge-based compute nodes as well as cloud resources for full \gls{MEC} experimentation.

@@

EdgeNet deviates from the approaches described above and represents a fully software-based, massively distributed public platform for Edge Computing research~\cite{cappos2018edgenet,senel2021edgenet1,senel2021edgenet2}.
It consists of a public, modified and customized Kubernetes~\cite{kubernetes} cluster.
The EdgeNet software is available for anyone to download;
individuals and organizations can quickly deploy self-hosted edge nodes which are then made available to anyone using the platform.
The end result is a flexible and scalable global edge cloud simplifying research and prototype development in edge and cloud computing.
A similar, closely-related project is KubeEdge~\cite{xiong2018extend}, a software framework for extending Kubernetes clusters from the cloud to the edge.

Finally, a number of other, small-scale approaches also exist.
In~\cite{gedawy2016cumulus}, the authors propose Cumulus, a prototype testbed for edge compute offloading in \gls{IOT} with the explicit goal of providing a generic and heterogeneous platform for Edge Computing research.
The authors of~\cite{rimal2018experimental} propose an experimental testbed for two-tiered edge architectures in the context of \gls{FIWI} Edge Computing.
This testbed is then employed to validate wireless access and compute resource scheduling strategies in these systems.
In~\cite{yamanaka2021design}, the authors present an experimental Edge Computing testbed capable of determining the offload location of a workload based on desired latency parameters.
\cite{diao2019scalable} propose yet another Kubernetes-based testbed based for edge task offloading based on heterogeneous hardware including low-power single-board computers such as Raspberry Pis and Jetson Nanos.
Finally,\ \cite{moorthy2022cloudraft} proposes \gls{CLOUDRAFT}, a cloud-based framework for mobile network experimentation, with a focus on simplifying the management of testbed resources.
The goal of this project is to integrate, coordinate, share, and improve upon existing testbeds through a common interface.

>>>>

\todo[inline]{Still needs some work, don't review yet.}

@@

This dissertation complements existing work on benchmarking distributed system architectures.
A rich body of literature exists with examples of such approaches for Cloud computing.
However, existing approaches tend to focus heavily on large-scale data processing and downlink transmission in Cloud scenarios, and are not applicable to the latency-sensitive, uplink-heavy characteristics of Edge Computing.
For example, the \gls{YCSB}~\cite{cooper2010benchmarking} is a data-intensive benchmark for the profiling of data centers and Cloud services; workload characterization in datacenters leads to a similar benchmark in~\cite{jia2013characterizing}.
In~\cite{turner2012cmart}, the authors introduce \emph{C-MART}, a benchmark for modern web applications running on the Cloud.
The benchmark utilizes modern web technologies, can be deployed automatically at a large scale, and includes generators that mimic remote clients accessing the system over the internet.\@
\cite{malawski2018benchmarking,back2018using} focus instead on serverless computing, using compute-intensive workloads to benchmark major Cloud function providers such as \gls{AWS}, Microsoft Azure, and \gls{GCP}.

The limitations of previous approaches has in later years led to a number of works that have attempted to tackle this challenge in the context of Fog and Edge Computing as well.
Even so, these works have generally leveraged static workloads that fail to account for the intensely dynamic nature of \gls{CPS} deployments.
Furthermore, none of these approaches have attempted to understand the relationship between latency and human performance in \gls{XR} applications deployed on the Edge.
\emph{EdgeBench}, introduced in~\cite{das2018edgebench}, corresponds to a benchmarking suite for the Edge.
It emulates three different workloads, speech-to-text, image recognition, and sensor network application traffic, and specifically targets serverless deployments.\@
\cite{baurle2022comb} presents \emph{ComB}, a flexible and extensible, application-oriented benchmarking suite for Edge computing.
It includes a built-in distributed video-analytics benchmark, but is designed to decouple orchestration and workload, allowing for easy extension of the latter.\@
\emph{IoTBench}~\cite{lee2019iotbench} is a benchmarking tool for Edge-deployed \gls{IOT} applications.
It includes seven representative workloads, and allows researchers to evaluate the performance of these applications on Edge infrastructure through the analysis of computational demands and resource consumption.\@
\cite{mcchesney2019defog} presents a benchmarking suite for the comparison of deployments on the Edge, Fog, or Cloud named \emph{DeFog}.
\emph{DeFog} includes six latency-critical and bandwidth-intensive, Fog- and Edge-relevant workloads, including real-time video detection and a text-to-speech application.
In~\cite{lockhart2020scission}, the authors introduce a tool, \emph{Scission}, for the benchmarking of \glspl{DNN} partitioned across the user device, Edge, and Cloud continuum.
The tool can also be used to automatically determine the optimal partition of a \gls{DNN}.
Finally, \emph{OpenRTiST}~\cite{george2020openrtist} is a tool for the benchmarking and performance evaluation of Edge Computing deployments using real-time processing of a live video feed.
Video is streamed from a source Android device to a compute node where neural style-transfer~\cite{gatys2016image} is applied to it.
The transformed video is then fed back to the source to be presented, resulting in a compute-intensive, bandwidth-hungry, and latency-sensitive benchmark.

@@

\gls{GENI} is a testbed deployed across more than fifty sites across the \gls{USA}.
Originally designed for ``the development, deployment, and validation of transformative, at-scale concepts in network science, services, and security''~\cite{berman2014geni}, \gls{GENI} today combines local Edge compute resources with mobile network last-hop connectivity for \gls{MEC} experimentation~\cite{gosain2017geni}.
The wireless-capable sites incorporate single-hop access to compute, storage, and network resource, making \gls{GENI} a compelling Edge Computing platform.

The Chameleon testbed is a research platform originally designed for the study and evaluation of Cloud computing systems~\cite{keahey2020lessons}, but which has recently been expanded to target Edge Computing research through the \acs{CHI}@Edge~\cite{chiatedge} project.
Chameleon is a collaborative project operated by a consortium of research institutions and industry partners, and is designed to provide a flexible and reconfigurable infrastructure for researchers to experiment with and test various Cloud computing technologies, algorithms, and configurations.
The testbed provides a highly reconfigurable array of hardware and software resources, including servers, storage systems, and networking equipment, able to simulate various Cloud and Edge Computing scenarios.

\gls{COSMOS} is a testbed deployed in New York City (State of New York, \gls{USA}), containing over \num{200} rooftop, intermediate, and mobile nodes.
The testbeds main \emph{raison d'être} is \gls{MEC} research, with a strong emphasis on wireless networks.
The system includes \glspl{SDR}, \si{\milli\meter}-wave equipment, optical fibers, Cloud integration, and Edge compute for core network functionality and application data processing~\cite{yu2019cosmos,raychaudhuri2020challenge}.

\gls{POWDER} is described by the authors of~\cite{breen2020powder} as a ``city-scale, remotely accessible, end-to-end software defined platform supporting a broad range of wireless and mobile related research''.
It features roughly \num{15} fixed programmable radio nodes, based on off-the-shelves \glspl{SDR}, distributed across a \SI{15}{\kilo\meter\squared} area in Salt Lake City (State of Utah, \gls{USA}).
These radio nodes are combined with Edge-based compute nodes as well as Cloud resources for full \gls{MEC} experimentation.

@@

EdgeNet deviates from the approaches described above and represents a fully software-based, massively distributed public platform for Edge Computing research~\cite{cappos2018edgenet,senel2021edgenet1,senel2021edgenet2}.
It consists of a public, modified and customized Kubernetes~\cite{kubernetes} cluster.
The EdgeNet software is available for anyone to download;
individuals and organizations can quickly deploy self-hosted Edge nodes which are then made available to anyone using the platform.
The end result is a flexible and scalable global Edge Cloud simplifying research and prototype development in Edge and Cloud computing.
A similar, closely-related project is KubeEdge~\cite{xiong2018extend}, a software framework for extending Kubernetes clusters from the Cloud to the Edge.

Finally, a number of other, small-scale approaches also exist.
In~\cite{gedawy2016cumulus}, the authors propose Cumulus, a prototype testbed for Edge compute offloading in \gls{IOT} with the explicit goal of providing a generic and heterogeneous platform for Edge Computing research.
The authors of~\cite{rimal2018experimental} propose an experimental testbed for two-tiered Edge architectures in the context of \gls{FIWI} Edge Computing.
This testbed is then employed to validate wireless access and compute resource scheduling strategies in these systems.
In~\cite{yamanaka2021design}, the authors present an experimental Edge Computing testbed capable of determining the offload location of a workload based on desired latency parameters.
\cite{diao2019scalable} propose yet another Kubernetes-based testbed based for Edge task offloading based on heterogeneous hardware including low-power single-board computers such as Raspberry Pis and Jetson Nanos.
Finally,\ \cite{moorthy2022cloudraft} proposes \gls{CLOUDRAFT}, a Cloud-based framework for mobile network experimentation, with a focus on simplifying the management of testbed resources.
The goal of this project is to integrate, coordinate, share, and improve upon existing testbeds through a common interface.
